---
title: "Untitled"
output: html_document
---

```{r}
```{r}
# All the functions and libraries needed for this script
library(boot)
library(limma)
library(pheatmap)
library(RColorBrewer)
library(viridis)
library(dplyr)
library(gplots)
library(heatmap3)
library(caret)
library(nnet)
library(ROCR)
library(ranger)
library(data.table)
library(xgboost)
library(Matrix)
library(glmnet)
library(pROC)
library(mRMRe)
library(tibble)
library(SummarizedExperiment)
library(edgeR)
library(affy)

make_label <- function(HTSeq_matrix) {
  label_HTSeq <- factor(c(0, 1))
  x = 1
  for (i in rownames(HTSeq_matrix)) {
    if (substr(i, start = 1, stop = 4) == "LUAD") {
      label_HTSeq[x] = 1
    } else {
      label_HTSeq[x] = 0
    }
    x = x + 1
  }
  return (label_HTSeq)
}
mean_squared_error <- function(data, index) {
  responses <- data$response[index]
  predictions <- data$prediction[index]
  MSE <- mean((responses - predictions)^2)
  return(MSE)
}
mean_absolute_error <- function(data, index) {
  responses <- data$response[index]
  predictions <- data$prediction[index]
  MAE <- mean(abs(responses - predictions))
  return(MAE)
}
OS_R_squared <- function(data, index) {
  responses <- data$response[index]
  predictions <- data$prediction[index]
  baseline <- data$baseline[index]
  SSE <- sum((responses - predictions)^2)
  SST <- sum((responses - baseline)^2)
  r2 <- 1 - SSE/SST
  return(r2)
}
OSR2 <- function(prediction, test, train) {
  SSE = sum((test - prediction)^2)
  SST = sum((test - mean(train))^2)
  OSR2 = 1 - SSE/SST
  return (OSR2)
}
tableAccuracy <- function(test, pred) {
  t = table(test, pred)
  a = sum(diag(t)/length(test))
  return(a)
}

accuracy <- function(data, index) {
  responses <- data$response[index]
  predictions <- data$prediction[index]
  
  acc <- tableAccuracy(responses, predictions)
  return (acc)
  
}
all_metrics <- function(data, index) {
  mse <- mean_squared_error(data, index)
  mae <- mean_absolute_error(data, index)
  OSR2 <- OS_R_squared(data, index)
  accurate <- accuracy(data, index)
  return(c(mse, mae, OSR2, accurate))
}

reverse_log <- function(data) {
  return (2^data)
}
```


```{r}

curated <- read.csv("curated_dataset.csv", header = FALSE)

head(curated)


curated[,1]


bladder_cancer <- data.frame(1:nrow(curated))
hep_cancer <- data.frame(1:nrow(curated))
bre_cancer <- data.frame(1:nrow(curated))
crc_cancer <- data.frame(1:nrow(curated))
eso_cancer <- data.frame(1:nrow(curated))
gas_cancer <- data.frame(1:nrow(curated))
lung_cancer <- data.frame(1:nrow(curated))
ova_cancer <- data.frame(1:nrow(curated))
glio_cancer <- data.frame(1:nrow(curated))
sar_cancer <- data.frame(1:nrow(curated))
pan_cancer <- data.frame(1:nrow(curated))
pro_cancer <- data.frame(1:nrow(curated))
bil_cancer <- data.frame(1:nrow(curated))
non_cancer <- data.frame(1:nrow(curated))

bla = 1
hep = 1
bre = 1
crc = 1
eso = 1
gas = 1
lung = 1
ova = 1
glio = 1
sar = 1
pan = 1
pro = 1
bil = 1
non = 1

for (i in 1:(length(colnames(curated)) - 1)) {
  cancer_type <- curated[,i+1][1]
  cancer_type <- substr(cancer_type, start = 1, stop = 2)
  print(i)
  print(cancer_type)
  if (cancer_type == "Bl") {
    name <- paste("Bladder_", toString(bla))
    bladder_cancer[name] <- curated[,i+1]
    bla = bla + 1
  } else if (cancer_type == "He" | cancer_type == "HC"){
    name <- paste("HCC_", toString(hep))
    hep_cancer[name] <- curated[,i+1]
    hep = hep + 1
  } else if (cancer_type == "Br") {
    name <- paste("Breast_", toString(bre))
    bre_cancer[name] <- curated[,i+1]
    bre = bre + 1 
  } else if (cancer_type == "Co" | cancer_type == "CC") {
    name <- paste("CRC_", toString(crc))
    crc_cancer[name] <- curated[,i+1]
    crc = crc + 1 
  } else if (cancer_type == "Es" | cancer_type == "EC") {
    name <- paste("Esophageal_", toString(eso))
    eso_cancer[name] <- curated[,i+1]
    eso = eso + 1 
  } else if (cancer_type == "Ga" | cancer_type == "GC") {
    name <- paste("Gastric_", toString(gas))
    gas_cancer[name] <- curated[,i+1]
    gas = gas + 1 
    
  } else if (cancer_type == "Lu" | cancer_type == "Lk") {
    name <- paste("Lung_", toString(lung))
    lung_cancer[name] <- curated[,i+1]
    lung = lung + 1 
    
  } else if (cancer_type == "Ov") {
    name <- paste("Ovarian_", toString(ova))
    ova_cancer[name] <- curated[,i+1]
    ova = ova + 1 
    
  } else if (cancer_type == "GL" | cancer_type == "Gl") {
    name <- paste("Glioma_", toString(glio))
    glio_cancer[name] <- curated[,i+1]
    glio = glio + 1 
  } else if (cancer_type == "SA" | cancer_type == "Sa") {
    name <- paste("Sarcoma_", toString(sar))
    sar_cancer[name] <- curated[,i+1]
    sar = sar + 1 
  } else if (cancer_type == "Pa" | cancer_type == "PC") {
    name <- paste("Pancreatic_", toString(pan))
    pan_cancer[name] <- curated[,i+1]
    pan = pan + 1 
  } else if (cancer_type == "PR") {
    name <- paste("Prostate_", toString(pro))
    pro_cancer[name] <- curated[,i+1]
    pro = pro + 1 
    
  } else if (cancer_type == "Bi" | cancer_type == "BT") {
    name <- paste("Biliary_", toString(bil))
    bil_cancer[name] <- curated[,i+1]
    bil = bil + 1 
  } else{
    name <- paste("NonCancer_", toString(non))
    non_cancer[name] <- curated[,i+1]
    non = non + 1 
  }
}


#Randomly divide into Discovery and Validation set for each cancer

set.seed(99)
bladder_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(bladder_cancer), floor(0.7*ncol(bladder_cancer)))
bladder.train <- bladder_cancer[,disc]
bladder.test <- bladder_cancer[,-disc]

set.seed(101)
hep_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(hep_cancer), floor(0.7*ncol(hep_cancer)))
hep.train <- hep_cancer[,disc]
hep.test <- hep_cancer[,-disc]

set.seed(103)
bre_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(bre_cancer), floor(0.8*ncol(bre_cancer)))
bre.train <- bre_cancer[,disc]
bre.test <- bre_cancer[,-disc]


set.seed(105)
crc_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(crc_cancer), floor(0.8*ncol(crc_cancer)))
crc.train <- crc_cancer[,disc]
crc.test <- crc_cancer[,-disc]

set.seed(107)
eso_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(eso_cancer), floor(0.8*ncol(eso_cancer)))
eso.train <- eso_cancer[,disc]
eso.test <- eso_cancer[,-disc]

set.seed(109)
gas_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(gas_cancer), floor(0.8*ncol(gas_cancer)))
gas.train <- gas_cancer[,disc]
gas.test <- gas_cancer[,-disc]

set.seed(111)
lung_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(lung_cancer), floor(0.8*ncol(lung_cancer)))
lung.train <- lung_cancer[,disc]
lung.test <- lung_cancer[,-disc]

set.seed(113)
ova_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(ova_cancer), floor(0.7*ncol(ova_cancer)))
ova.train <- ova_cancer[,disc]
ova.test <- ova_cancer[,-disc]

set.seed(115)
glio_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(glio_cancer), floor(0.8*ncol(glio_cancer)))
glio.train <- glio_cancer[,disc]
glio.test <- glio_cancer[,-disc]

set.seed(117)
sar_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(sar_cancer), floor(0.8*ncol(sar_cancer)))
sar.train <- sar_cancer[,disc]
sar.test <- sar_cancer[,-disc]

set.seed(119)
pan_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(pan_cancer), floor(0.8*ncol(pan_cancer)))
pan.train <- pan_cancer[,disc]
pan.test <- pan_cancer[,-disc]

set.seed(121)
pro_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(pro_cancer), floor(0.4*ncol(pro_cancer)))
pro.train <- pro_cancer[,disc]
pro.test <- pro_cancer[,-disc]

set.seed(107)
bil_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(bil_cancer), floor(0.8*ncol(bil_cancer)))
bil.train <- bil_cancer[,disc]
bil.test <- bil_cancer[,-disc]

set.seed(123)
non_cancer$X1.nrow.curated. <- NULL
disc <-  sample(ncol(non_cancer), floor(0.8*ncol(non_cancer)))
non.train <- non_cancer[,disc]
non.test <- non_cancer[,-disc]


Discovery_set <- Reduce(cbind, list(bladder.train, hep.train, crc.train, bre.train, pan.train, glio.train, sar.train, lung.train, ova.train, eso.train, gas.train, pro.train, bil.train, non.train))
Test_set <- Reduce(cbind, list(bladder.test, hep.test, crc.test, bre.test, pan.test, glio.test, sar.test, lung.test, ova.test, eso.test, gas.test, pro.test, bil.test, non.test))


#write.csv(Discovery_set, "Discovery_set.csv", sep='\t')
#write.csv(Test_set, "Test_set.csv", sep='\t')

```

```{r}

# Doing Welch's t test and generating FDR and filtering top 500 miRNAs and writing in the table as "Curated_training_genes.csv" to be used later
rowname <- row.names(data)
data <- as.data.frame(t(data), )
data <- tibble::rownames_to_column(data, "Samples")

head(data)

union_genes <- subset(data, Samples %in% gene_frequency$miRNA[1:5])

union_genes <- as.data.frame(t(union_genes), )

colnames(union_genes) <- union_genes[1,]
union_genes <- union_genes[-1, ] 


t_test_lst <- vector(mode = "list")
p_values_lst <- 1:length(colnames(union_genes))
logFC_lst <- 1:length(colnames(union_genes))
x = 1
for (i in colnames(union_genes)){
  if (i == "isCancer") {
    break
  }
  t_test_lst[[i]] <- t.test(as.numeric(union_genes[[i]][1:2250]), as.numeric(union_genes[[i]][2251:4497]), var.equal = FALSE)
  union_genes[[i]] <- as.numeric(union_genes[[i]])
  p_values_lst[x] <- t_test_lst[[i]]$p.value
  logFC <- log2(mean(union_genes[[i]][1:2250])/mean(union_genes[[i]][2251:4497]))
  logFC_lst[x] <- logFC
  x = x + 1
  
}

#write.csv(union_genes_data, "Curated_training_Genes.csv", sep='\t')
```


```{r}
# Now generating 100 random forest models

cancer <- read.csv("Discovery.csv", header = TRUE)
dge <- read.csv("Curated_training_Genes.csv", header = TRUE)


top_miRNA <- dge$data.Samples[1:500]
top_miRNA
head(cancer)
cancer <- subset(cancer,Gene %in% top_miRNA )
miRNA <- cancer$Gene
cancer$Gene <- NULL
row.names(cancer) <- miRNA

cancer <- 2^cancer

label <- factor(c("A", "B"))
column <- 1:length(colnames(cancer))
x = 1
j = 1
s = 1
for (i in colnames(cancer)) {
  if (substr(i, start = 1, stop = 3) == "Non") {
    label[x] = "B"
    column[x] <- paste("normal_", toString(j))
    j = j + 1
  } else {
    label[x] = "A"
    column[x] <- paste("cancer_", toString(s))
    s = s + 1
  }
  x = x + 1
}
colnames(cancer) <- column
data <- as.data.frame(t(cancer))
data$isCancer <- label


seed = 35 
ConfusionMatrices <- vector("list", 100)
for (i in 1:100) {
 ConfusionMatrices[[i]] <- matrix(NA, nrow=2, ncol=2)
}

for (i in 1:100) {
  set.seed(seed)
  train <-  sample(nrow(data), floor(0.7*nrow(data)))
  train.micro <- data[train,]
  test.micro <- data[-train,]
  
  tgrid <- expand.grid(
  .mtry = 1:30,
  .splitrule = "gini", 
  .min.node.size = 1
  )
  set.seed(seed + 11)
  rf.cv <- train(isCancer ~ ., 
                   data = train.micro,
                   method = "ranger",
                   tuneGrid = tgrid,
                   trControl = trainControl(method = "cv", number = 10, verboseIter = TRUE)
                   )
  mtry <- rf.cv$bestTune$mtry
  
  set.seed(seed + 11)
  rf_model <- ranger(isCancer ~ ., mtry = mtry, data = train.micro, importance = "impurity")
  rf_pred <- predict(rf_model, data = test.micro)
  ConfusionMatrices[[i]] <- table(rf_pred$predictions, test.micro$isCancer) 
  ranger_gini <- as.data.frame(rf_model$variable.importance)
  
  setDT(ranger_gini, keep.rownames = TRUE)[]
  colnames(ranger_gini) <- c("Gene", "Gini")
  ranger_gini_table <- ranger_gini %>% arrange(desc(Gini))
  gini_table_name <- paste("rf_gini_", toString(i))
  write.csv(ranger_gini_table, gini_table_name, sep='\t')
  seed = seed + 99
}
```

```{r}
# Getting the top 10 miRNAs from 100 random forest models
lst <- vector(mode = "list", length = 100)

for (i in 1:100) {
  file <- paste("rf_gini_", toString(i))
  gini <- read.csv(file, header = TRUE, sep = ",")
  lst[[i]] <- gini$Gene[1:10]
}


miRNA <- Reduce(union, lst)
frequency <- 1:length(miRNA)
k = 1
for (gene in miRNA) {
  count = 0
  for (i in 1:length(lst)) {
    if (gene %in% lst[[i]]) {
      count = count + 1
    } 
  }
  frequency[k] <- count 
  k = k + 1
}


gene_frequency <- data.frame(miRNA, frequency)

gene_frequency <- gene_frequency %>% arrange(desc(frequency))

gene_frequency$miRNA[1:5]
gene_frequency
```


```{r}

```{r}
# Testing validation set

valid <- read.csv("Validation_Set.csv", header = TRUE)

miRNA_valid <- valid$Gene 
valid$Gene <- NULL
row.names(valid) <- miRNA_valid

valid <- 2^valid

head(valid)

label <- factor(c(1, 0))
column <- 1:1664
x = 1
j = 1
s = 1
for (i in colnames(valid)) {
  if (substr(i, start = 1, stop = 3) == "Non") {
    label[x] = 0
    column[x] <- paste("normal_", toString(j))
    j = j + 1
  } else {
    label[x] = 1
    column[x] <- paste("cancer_", toString(s))
    s = s + 1
  }
  x = x + 1
}
colnames(valid) <- column
valid_data <- as.data.frame(t(valid))
valid_data$isCancer <- label


valid_data <- as.data.frame(t(valid_data), )
valid_data <- tibble::rownames_to_column(valid_data, "Samples")


valid_union_genes <- subset(valid_data, Samples %in% gene_frequency$miRNA[1:5])

valid_union_genes <- as.data.frame(t(valid_union_genes), )

colnames(valid_union_genes) <- valid_union_genes[1,]
valid_union_genes <- valid_union_genes[-1, ] 

t_test_lst <- vector(mode = "list")
p_values_lst <- vector(mode = "list")
for (i in colnames(valid_union_genes)){
  
  t_test_lst[[i]] <- t.test(as.numeric(valid_union_genes[[i]][1:1102]), as.numeric(valid_union_genes[[i]][1103:1664]), var.equal = FALSE)
  valid_union_genes[[i]] <- as.numeric(valid_union_genes[[i]])
  p_values_lst[[i]] <- t_test_lst[[i]]$p.value
  
}

t_test_lst
p_values_lst

cancer_label <- factor(c(1, 0))
y = 1 
for (i in row.names(valid_union_genes)) {
  if (substr(i, start = 1, stop = 3) == "can") {
    cancer_label[y] <- 1
  } else {
    cancer_label[y] <- 0
  }
  y = y + 1
}

valid_union_genes$isCancer <- cancer_label

```

```{r}
# AUC on Discovery set for reference (AUC on validation set done the same way but just using the same model built on discovery set on validation set)

MIMAT0003326_model <- multinom(isCancer ~ MIMAT0003326, family = "binomial", data = union_genes)
pred_MIMAT0003326 <- predict(MIMAT0003326_model, valid_union_genes, type = 'prob')
pred_MIMAT0003326 <- ROCR::prediction(pred_MIMAT0003326, valid_union_genes$isCancer)
roc_MIMAT0003326 <- ROCR::performance(pred_MIMAT0003326, "tpr", "fpr")

MIMAT0027504_model <- multinom(isCancer ~ MIMAT0027504, family = "binomial", data = union_genes)
pred_MIMAT0027504 <- predict(MIMAT0027504_model, valid_union_genes, type = 'prob')
pred_MIMAT0027504 <- ROCR::prediction(pred_MIMAT0027504, valid_union_genes$isCancer)
roc_MIMAT0027504 <- ROCR::performance(pred_MIMAT0027504, "tpr", "fpr")


MIMAT0027468_model <- multinom(isCancer ~ MIMAT0027468, family = "binomial", data = union_genes)
pred_MIMAT0027468 <- predict(MIMAT0027468_model, valid_union_genes, type = 'prob')
pred_MIMAT0027468 <- ROCR::prediction(pred_MIMAT0027468, valid_union_genes$isCancer)
roc_MIMAT0027468 <- ROCR::performance(pred_MIMAT0027468, "tpr", "fpr")

MIMAT0015064_model <- multinom(isCancer ~ MIMAT0015064, family = "binomial", data = union_genes)
pred_MIMAT0015064 <- predict(MIMAT0015064_model, valid_union_genes, type = 'prob')
pred_MIMAT0015064 <- ROCR::prediction(pred_MIMAT0015064, valid_union_genes$isCancer)
roc_MIMAT0015064 <- ROCR::performance(pred_MIMAT0015064, "tpr", "fpr")

MIMAT0031000_model <- multinom(isCancer ~ MIMAT0031000, family = "binomial", data = union_genes)
pred_MIMAT0031000 <- predict(MIMAT0031000_model, valid_union_genes, type = 'prob')
pred_MIMAT0031000 <- ROCR::prediction(pred_MIMAT0031000, valid_union_genes$isCancer)
roc_MIMAT0031000 <- ROCR::performance(pred_MIMAT0031000, "tpr", "fpr")


combined_model <- multinom(isCancer ~ MIMAT0003326 + MIMAT0027504  + MIMAT0015064 + MIMAT0031000, family = "binomial", data = union_genes) #all the combinations tested. Best one is this one with only 4 miRNA
pred_combined <- predict(combined_model, valid_union_genes, type = 'prob')
pred_combined <- ROCR::prediction(pred_combined, valid_union_genes$isCancer)
roc_combined <- ROCR::performance(pred_combined, "tpr", "fpr")

```

